---
title: "Gapminder!"
author: "Kaitlyn Harper"
date: "September 14, 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Hello! We're going to explore the Gapminder dataset today! First, we need to load the dataset from Rstudio. Note: for this markdown, you'll see my code in gold and the output in blue! 

<style>
div.color pre { background-color:lightskyblue; }
div.color pre.r { background-color:gold; }
</style>

<div class = "color">
```{r View_Load_Gapminder, echo = TRUE, background = "plum"}

# Load tidyverse + all corresponding packages!
library(tidyverse)

# Load gapminder dataset
library(gapminder)
data("gapminder")

# View the first 6 lines of the data set (to make sure it's there)
head(gapminder)

# This is also a great way to look at the head of this data set! 
glimpse(gapminder)

```
</div>
Sweet! So it looks like we have a few different types of variables in our data set. The `r names(gapminder)[1]` and `r names(gapminder)[2]` variables are **`r paste(class(gapminder$country), "s", sep = "")`**; the `r names(gapminder)[3]` and `r names(gapminder)[5]` variables are **`r paste(class(gapminder$year), "s", sep = "")`**; the `r names(gapminder)[4]` and `r names(gapminder)[6]` variables are **`r class(gapminder$lifeExp)`**.

Okie doke. Let's take a look at some of the variables now. For example, if we want to find out how many unique countries are in this data set, we could do this:
<div class = "color">
```{r unique_countries, echo=TRUE}

# Unique number of countries
length(unique(gapminder$country))

```
</div>
Looks like there are 142 countries in our data set. Interesting. I wonder if all the continents are present...
<div class = "color">
```{r continents, echo=TRUE}

unique(gapminder$continent)

```
</div>
Huh, looks like North and South America are lumped into one, and Antarctica is missing. 

I'm pretty sure we're going to get to plotting in the next hw assignment, so let's just explore some of the simple stats for this data set. 







